---
title: "Text Mining With R"
output: html_document
---

[Source : Text Mining With R](http://tidytextmining.com/tidytext.html#the-unnest_tokens-function)

## Chapter 1 : The tidy text format


```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
getwd()
```

### Package Management

```{r, message=FALSE, warning=FALSE}

library(dplyr)
library(tidytext)

```


### The unnest_tokens function

All text needs to be converted to tidy format using the unnest_tokens function.

Load some sample text

```{r}

text <- c("Because I could not stop for Death -",
          "He kindly stopped for me -",
          "The Carriage held but just Ourselves -",
          "and Immortality")

text
str(text)


```



Convert to data frame

```{r}

text_df<-data_frame(line=1:4,text=text)
text_df

```

Break into tokens now

```{r}
text_df %>% 
  unnest_tokens(word,text) # Output column, input col
  
```

This automatically removes punctuations, lowers the case etc.

### Tidying the works of Jane Austen

Load thw books 

```{r , warning=FALSE}

library(stringr)
library(janeaustenr)
library(dplyr)

original_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]",
                                                 ignore_case = TRUE)))) %>%
  ungroup()

original_books

  


```


DF is already in tidy form. 
Restructure using unnest function

```{r}

library(tidytext)

original_books %>% 
  unnest_tokens(word,text)->tidy_text

head(tidy_text)

```

















Each row in this df now has a word as a token.
Token units can be changed to sentenence etc.


Remove stop words using the built in stop words df and the anti join function

```{r}

data("stop_words") #comes with the package
head(stop_words)

tidy_text %>% 
  anti_join(stop_words)->tidy_text  

head(tidy_text)

```




Find most frequent words from this new DF
```{r}

tidy_text %>% 
  count(word,sort=TRUE)
```







Plot these words

```{r}

library(ggplot2)

tidy_text %>% 
  count(word,sort=TRUE) %>% 
  filter(n>600) %>% 
  mutate(word=reorder(word,n)) %>% 
  ggplot(aes(word,n))+
  geom_col()+
  coord_flip()


```


### Comparing Word frequencies across different novels

We find some novels of the same era as that of Jane Austen's and compare these works in terms of word freq.

Below novels are considered..

Author          Novel
--------        ----------
HG wells         The Time Machine
HG wells         The War of the Worlds
HG wells         The Invisible Man
HG wells         The Island of Doctor Moreau
Brontë sisters   Jane Eyre
Brontë sisters   Wuthering Heights 
Brontë sisters   The Tenant of Wildfell Hall 
Brontë sisters   Villette 
Brontë sisters   Agnes Grey



```{r}
library(gutenbergr)

hgwells <- gutenberg_download(c(35, 36, 5230, 159))


tidy_hgwells<-hgwells %>% 
  unnest_tokens(word,text) %>% 
  anti_join(stop_words)

head(tidy_hgwells)



```



Finding the most freq words in HG Wells novels

```{r}
tidy_hgwells %>% 
  count(word,sort=TRUE)

```


Same excercise with the Bronte sisters novels

```{r}


bronte <- gutenberg_download(c(1260, 768, 969, 9182, 767))

tidy_bronte <- bronte %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words)

tidy_bronte %>%
  count(word, sort = TRUE)
```



Lets do this excercise with all novels combined

```{r}
library(tidyr)

frequency<-bind_rows( mutate(tidy_text,author="Jane Austin"),
                      mutate(tidy_bronte,author="Bronte Sisters"),
                      mutate(tidy_hgwells,author="HG Wells")) %>% 
  mutate(word = str_extract(word, "[a-z']+")) %>%       ## To make _any_ and any similar. Issue with gutenberg encoding
  count(author, word)%>% 
  group_by(author) %>% 
  mutate(proportion=n/sum(n)) %>% 
  select(-n) %>% 
  spread(author,proportion) %>% 
  gather(author,proportion,`Bronte Sisters`:`HG Wells`)





frequency


```


Plot the proportion of the word freq with jane austin's word freq

```{r}

library(scales)

# expect a warning about rows with missing values being removed
ggplot(frequency, aes(x = proportion, y = `Jane Austin`, color = abs(`Jane Austin` - proportion))) +
  geom_abline(color = "gray40", lty = 2) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  scale_color_gradient(limits = c(0, 0.001), low = "darkslategray4", high = "gray75") +
  facet_wrap(~author, ncol = 2) +
  theme(legend.position="none") +
  labs(y = "Jane Austin", x = NULL)

```


Words that are close to the line in these plots have similar frequencies in both sets of texts
Words in the Austen-Brontë panel are closer to the zero-slope line than in the Austen-Wells panel
Austen and the Brontë sisters use more similar words than Austen and H.G. Wells
!