---
title: "K Fold Cross Validation with Caret"
author: "Rizwaan Adil"
date: "January 31, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***
>Objective : Demo caret package usage for K Fold Cross Validation

***
#### Package Management


<br>
Load required packages and check the data
```{r, message=FALSE}

library(caret)
library(kernlab)
data(spam)
head(spam)

```

Here the variable to be predicted is __type__ which identies a spam or ham message


***
#### Data Slicing Operations 
***

##### Splitting the data into test and training sets


- Splitting is done using the __CreateDataPartition__ Function
- It returns an index which is used to create test and training sets

```{r, message=FALSE}
dim(spam)
trainIndex<-createDataPartition(y = spam$type,p=0.75,list=FALSE,times = 1) # Use the type column ; 75% data goes to training, return a matrix
head(trainIndex)



```

An index with 4601 rows (same as original dataset) is returned with sampled row names

<br>

Create test and training sets using the index

```{r}
training<-spam[trainIndex,]
dim(training)

test<-spam[-trainIndex,]
dim(test)


```

Rows in training set:  3451
Rows in test set : 1150


##### K Folds

<br> 
- Used for performing K fold cross validation
- Done using __createFolds__ function

```{r, message=FALSE}

set.seed(123)

folds<-createFolds(y=spam$type,k=10,list = TRUE,returnTrain = FALSE) # Return a list of folds , return only the test set

sapply(folds,length)
```

<br>
Review the individual elements in a  fold

```{r}
folds$Fold01[1:10]

```



<br>
##### Resampling for Bootrstrapping 
<br>

```{r}

set.seed(123)

samples<-createResample(y=spam$type,times = 10,list = TRUE) # Return a list of 10 resamples ; with replacement

sapply(samples,length)


```


<br>
Accessing individual elements

```{r}
samples$Resample01[1:10]
```










####__Resources__
- [Practical Machine Learning](https://www.coursera.org/learn/practical-machine-learning/lecture/0vwNS/training-option)
- [Resampling Methods](https://lagunita.stanford.edu/courses/HumanitiesSciences/StatLearning/Winter2016/courseware/9956347366744e1cac95f513e9235f9f/9339db132f504f809056f221271d9795/)
