---
title: "Week 1 - Building Data Visualization Tools"
author: "Rizwaan Adil"
date: "January 3, 2017"
output: html_document

---

[Mooc Link](https://www.coursera.org/learn/r-data-visualization/supplement/1FG1D/initializing-a-ggplot-object)  
[Referenced Book](https://bookdown.org/rdpeng/RProgDA/basic-plotting-with-ggplot2.html#initializing-a-ggplot-object)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#### Package Management

Install necessary packages 

```{r , warning=FALSE, message=FALSE}
library("magrittr", lib.loc="~/R/win-library/3.3") ## Need to explicitly load it due to version conflict
library(ggplot2)
library(titanic)
library(faraway)
library(dplyr)

```

***

#### Dataset prep

```{r}

data("titanic_train")
titanic<-titanic_train
data("worldcup")
world_cup<-worldcup

names(titanic)

```

***

#### ggplot Modus Operandi


The basic steps behind creating a plot with ggplot2 are:  

1. Create an object of the ggplot class, typically specifying the data and some or all of the aesthetics;  
2. Add on geoms and other elements to create and customize the plot, using +  

__Aesthetics__  
1. Properties of the plots which are mapped to the variables to actually create the plot
2. Which aesthetics are required for a plot depend on which geoms you’re adding to the plot( try ?geom_point) 


***


### Basic Plotting

***


#### Creating a basic plot

Notes: 
 1. Do not start a line with the + sign
 2. Always use the aes() call to specify the properties\aesthetics of the plots. It is not recommended to do it anywhere else (possible but not recommended)

```{r}



titanic %>%
  ggplot()+
  geom_histogram(aes(x=Fare))


titanic %>%
  ggplot(aes(x=Fare))+
  geom_histogram()



```


***

#### Geoms

Notes

 1. Functions that add graphical elements to the plot
 2. Every function differs in arguments
 

```{r, warning=FALSE}

titanic %>% 
  ggplot(aes(x=Fare))+
  geom_histogram(bins = 15)

```


__Scatterplot__

```{r, warning=FALSE}

world_cup %>% 
  ggplot(aes(x=Time, y=Passes, color=Position, size= Shots ))+
  geom_point()

```

***

#### Multiple Geoms

Use multiple geoms to create several layers of info


__Ex.1__ : Adding point labels to a scatterplot

```{r, warning=FALSE}

##head(world_cup)

world_cup %>% 
  filter(Shots==max(Shots) | Passes==max(Passes)) %>% 
  mutate(pt_label=paste(Team,Position, sep=" , "))->gp



world_cup %>% 
  ggplot(aes(x=Passes,y=Shots))+
  geom_point()+
  geom_text(data=gp,aes(label=pt_label),vjust="inward",hjust="inward")



```


__Ex.2__ : Adding intercept lines


```{r, warning=FALSE}

world_cup %>% 
  ggplot(aes(x=Time))+
  geom_histogram(binwidth = 10)+
  geom_vline(xintercept = 90*0:6,
             color="red", alpha=0.9)


```

***

#### Extensions to GGPlot



Data prep

```{r, warning=FALSE}
nepali <- nepali %>%
  select(id, sex, wt, ht, age) %>%
  mutate(id = factor(id),
         sex = factor(sex, levels = c(1, 2),
                      labels = c("Male", "Female"))) %>%
  distinct(id, .keep_all = TRUE)

head(nepali)

```


Pair Plot


```{r, warning=FALSE, message=FALSE}

library(GGally)

ggpairs(nepali %>% select(sex, wt, ht,age))

```

***

### Customizing ggplot2 Plots

***

#### Guidlines for Good Plots

>1. Aim for high data density.  
>2. Use clear, meaningful labels.  
>3. Provide useful references.  
>4. Highlight interesting aspects of the data.  
>5. Consider using small multiples.  
>6. Make order meaningful.  



Data Prep

```{r , message=FALSE}

library(dplyr)
library(ggplot2)
library(gridExtra)
library(ggthemes)
library(dlnm)
library(faraway)


data(nepali)
data(worldcup)
data(chicagoNMMAPS)
chic <- chicagoNMMAPS
chic_july <- chic %>%
  filter(month == 7 & year == 1995)


```


 >High data density
 
 ***

If an element of the plot is redundant, take it out.
You should try to increase, as much as possible, the data to ink ratio in your graphs



 
 
```{r, message=FALSE}

world_cup %>% 
  ggplot(aes(x=Time, y= Shots))+
  geom_point()+
  theme_classic()




world_cup %>% 
  ggplot(aes(x=Time, y= Shots))+
  geom_point()+
  theme_tufte()

```





 > Meaningful Labels
 
 ***
 
 Do not use obscure acronyms
 
 
 
 
 
 