---
title: "Week 1 - Building Data Visualization Tools"
author: "Rizwaan Adil"
date: "January 3, 2017"
output: html_document

---

[Mooc Link](https://www.coursera.org/learn/r-data-visualization/supplement/1FG1D/initializing-a-ggplot-object)  
[Referenced Book](https://bookdown.org/rdpeng/RProgDA/basic-plotting-with-ggplot2.html#initializing-a-ggplot-object)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Package Management

Install necessary packages 

```{r , warning=FALSE}
library("magrittr", lib.loc="~/R/win-library/3.3") ## Need to explicitly load it due to version conflict
library(ggplot2)
library(titanic)
library(faraway)

```

***

#### Dataset prep

```{r}

data("titanic_train")
titanic<-titanic_train
data("worldcup")
world_cup<-worldcup

names(titanic)

```

***

#### ggplot Modus Operandi


The basic steps behind creating a plot with ggplot2 are:  

1. Create an object of the ggplot class, typically specifying the data and some or all of the aesthetics;  
2. Add on geoms and other elements to create and customize the plot, using +  

__Aesthetics__  
1. Properties of the plots which are mapped to the variables to actually create the plot
2. Which aesthetics are required for a plot depend on which geoms you’re adding to the plot( try ?geom_point) 


***

#### Creating a basic plot

Note: Do not start a line with the + sign

```{r}



titanic %>%
  ggplot()+
  geom_histogram(aes(x=Fare))

```



