---
title: "0. Customer Segmentation with R"
author: "Rizwaan Adil"
date: "January 3, 2017"
output: html_document

---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


***
Project Description :Segmentation using K Means and PCA  
Source              : [Rpubs](https://rpubs.com/hrbrmstr/customer-segmentation-r) |


#### Package Management
***

```{r , message=FALSE}

library(readxl)    # free data from excel hades
library(dplyr)     # sane data manipulation
library(tidyr)     # sane data munging
library(viridis)   # sane colors
library(ggplot2)   # needs no introduction
library(ggfortify) # super-helpful for plotting non-"standard" stats objects

```

***

#### Data Prep
***

Download data from url

```{r}

url<-"http://blog.yhathq.com/static/misc/data/WineKMC.xlsx"

filename<-basename(url)

if (!file.exists(filename))   download.file(url, filename)


```

  
<br>
Read in the file and rename columns

```{r}

offers<-read_excel(filename,sheet=1)
colnames(offers)

colnames(offers) <- c("offer_id", "campaign", "varietal", "min_qty", "discount", "origin", "past_peak")

colnames(offers)

head(offers)
```




Second dataset


```{r}

transactions <- read_excel(filename, sheet = 2)
colnames(transactions) <- c("customer_name", "offer_id")
transactions$n <- 1

head(transactions)

```

<br>
Join and transform data to wide form

```{r}

left_join(offers,transactions,by="offer_id") %>%  #join the tables
  count(customer_name,offer_id,wt=n) %>%               #group by customer and offer
  spread(offer_id,nn) %>%                              #pivot
  mutate_each(funs(ifelse(is.na(.),0,.))) -> raw

raw


```
***

#### Clustering

<br>
Using K - Means


```{r}

fit<-kmeans(raw[,-1],centers = 5,iter.max = 1000)
table(fit$cluster)

```
 


Plot the cluster info
```{r}


barplot(table(fit$cluster),col = "steelblue4")
```


***
#### Visualization of Clusters

Group customers by the offers they responded to..

```{r}

pca <- prcomp(raw[,-1])

pca_dat <- mutate(fortify(pca), col=fit$cluster) ## Consilidates the PCA loadings and cluster info for easier plotting

ggplot(pca_dat) +
  geom_point(aes(x=PC1, y=PC2, fill=factor(col)), size=3, col="#7f7f7f", shape=21) +
  scale_fill_viridis(name="Cluster", discrete=TRUE) + theme_bw()
```

Same visualization with autoplot function for clarity

```{r}

autoplot(fit,data=raw[,-1],frame=TRUE,frame.type="norm")

autoplot(fit,data=raw[,-1],frame=TRUE,frame.type="norm")

```


***
#### Exploring the original data with cluster info

Now that we have the clusters.. we might want to explore the data by cluster id.  
This way we can subset the original data and find interesting subgroups..

e.g.. lets pick out customers falling in the cluster no 4 and relevant details




```{r}


transactions %>% 
  left_join(data_frame(customer_name=raw$customer_name,cluster=fit$cluster)) %>% 
  left_join(offers) -> customer_clusters  ## give all the customers and clusters


## pick out cluster no 4 

customer_clusters %>% 
  mutate(is_4=(cluster==4)) %>% 
  count(is_4,varietal)
  

```

***

#### Summary

- PCA and Kmeans was used to cluster the customers by their response to marketing campaigns
